---
title: "example-breastcancer_cfDNA"
output: rmarkdown::html_vignette
Imports: tidyverse
vignette: >
  %\VignetteIndexEntry{example-breastcancer_cfDNA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Background

To test the `methyldeconvolveR` package, we used publicly available liquid biopsy data (methylated cfDNA) from [Hai et al. 2022 MedComm](https://doi.org/10.1002/mco2.134). In short, liquid biopsies were collected from patients with various breast cancer subtypes. Data was openly available for download from the NCBI SRA database, [SRP167041](https://trace.ncbi.nlm.nih.gov/Traces/?view=study&acc=SRP167041).

We curated a set of reference data, as described in [McNamara et al. 2023, JCI Insight](https://doi.org/10.1172/jci.insight.156529), and learned reference parameters (see the "Learn Reference" vignette), which we use for deconvolution. The reference atlas included as part of the package include all cell types described in the manuscript with the top 100 markers.

### Rationale

In this vignette, we show how to use `methyldeconvolveR` to perform cell type 
deconvolution from liquid biopsy samples. First, we load the relevant packages.

```{r setup}
library(methyldeconvolveR)
```

Next we load the relevant data. The `methyldeconvolveR` package comes with a
usable reference atlas to perform deconvolution in 15 different cell types.

We also read in the PAT file here. As part of the read-in, some (or likely many) reads
will be filtered out, which is reported when verbose = TRUE.

```{r load_data}
# First, load the generated or pre-existing reference
ref <- use_reference()

# Next, load the example sample PAT file
sample_pat_path <- system.file(package = "methyldeconvolveR", "extdata", "SRR8117413_dedup_sorted.pat.gz")
sample_pat <- read_pat(sample_pat_path, verbose = T)
head(sample_pat) # glimpse into the PAT file

```

### Analysis

Now, we can perform cell-type deconvolution in a single step with the `deconvolute_sample` function.

```{r deconv}
# NOTE: The path to the sample pat file or the pat data.frame itself can both be
# passed to the deconvolute_sample() function. Here, we pass the file path, 
# where read_pat is called internally.

results <- deconvolute_sample(sample_pat = sample_pat, reference = ref)
```

By default, 10 random initializations are used, and the log-likelihoods (LL) are computed.

In `results$best_result`, the initialization resulting in the maximum LL is returned.

#### Visualization

We can visualize the results here using `plot_deconv` function, which is built
using the `ggplot2` package.

```{r visualize}
plot_deconv(results)
```

